Understanding the impact of electronically-delivered parenting advice

### 2019-01-09 paper/cogsci2019/cogsci2019.tex
ported TeX draft to Rmd

---
title: "Understanding the impact of electronically-delivered parenting advice"
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{{\large \bf Hanwen Vivian~Zhang (vivian3@stanford.edu)} \\ Department of Psychology, Stanford University \\ Palo Alto, CA 94301 USA
    \AND {\large \bf George~Kachergis (george.kachergis@gmail.com)} \\ \\ Department of Psychology, Stanford University \\ Palo Alto, CA 94301 USA
    \AND {\large \bf Michael C.~Frank (mcfrank@stanford.edu)} \\ \\ Department of Psychology, Stanford University \\ Palo Alto, CA 94301 USA}

abstract: >
    Early parenting practices play an important role in shaping the future 
    outcomes of young children [@Hart1995; @Heckman2006]. 
    In particular, high quality early interactions and early language input appear to 
    facilitate more effective language learning and higher levels of school performance. 
    The rise in electronic parenting applications ("apps") holds the promise of delivering 
    low-cost, positive interventions on parenting style to a variety of different populations. 
    Of special interest are the parents of very young children, who are often difficult to reach in other ways. 
    Yet little is known about the effects of communicating to parents through app-based interventions.
    MTLD, types and tokens, joint attention, proportion of nouns/verbs/adjectives, 
    affect, concreteness, mean length of utterance...

keywords: >
    digital parenting advice; joint action; lexical diversity; guided play;

output: cogsci2016::cogsci_paper
#final-submission: \cogscifinalcopy
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, 
                      fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, 
                      message=F, sanitize = T)
```

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
```

# Introduction

- children learn through play -- physics, object features

## Guided Play Scaffolds Learning
- guided play activities can scaffold children's language and reasoning

## Improving Parenting Practices and Language Use
- early interventions on parenting practices can help children's outcomes

## Effectiveness of Digital Delivery
apps are an easy way to deliver curated, age-appropriate parenting advice — but can we measure an influence on parent's language? and do the interventions lead to children paying more attention?

@@

# Experiment 1

@@

## Method

### Participants. 
60 infants aged 6-24 months and their parents participated in a museum in northern California. We included infants who were exposed to English at least 50 percent of the time or who were exposed less but whose participating parent reported that they primarily speak English with their child at home.
[include specific demo info here?] - good to include now, and we can comment it out if it's too much (the full report/paper will have it)

### Materials.
Stimuli included videos from a commercial parenting application, Kinedu Inc. The videos were designed to show activities parents could perform with their child to foster cognitive and physical development, and were targeted to the child's age and development. In each video, an adult and child perform the activity while a narrator explains the activity and its purpose. We selected two videos for each of three age groups in our sample (6-11.9 months, 12-17.9 months, 18-23.94 months). More information about the specific videos is available in the Appendix. Participants were also given a set of props corresponding to those in the video they watched, so that they could complete the activity. The props associated with each video are listed in the Appendix.

@@

## Procedure.
After providing informed consent, parents in the Activity Video condition watched the assigned video on a laptop with headphones.
To ensure that parents could give the video their full attention, the experimenter played with the infant with a set of toys (different from the experimental props used in the study) while the video was being played. 
Immediately following the video, each parent-child dyad was provided with the props to complete the activity they had viewed. 
The toys were placed on a large foam mat, and parents were instructed to sit on the mat with their child and re-create the activity they had viewed for a period of three minutes.
In the No-Video control condition, after consenting parents were told to play with their child as they would at home with the provided props for a period of three minutes. 
They were not given any additional instructions about how to use the props.

@@

### Joint Attention Coding Procedure.
Video of each session was coded for episodes of joint attention using Datavyu software. 
The video taken at floor level was coded by default, but the other video was referred to if the participants were not visible or if there was a technological error with the first camera. For each participant we coded for episodes of coordinated joint attention, episodes of passive joint attention, and parental bids for joint attention. 
We defined parental bids for joint attention as any attempt to initiate joint attention (i.e label, point, or otherwise draw attention to an object) that does not result in passive or coordinated joint attention. 
If more than 3 seconds elapsed between bids, they were coded as separate attempts. 
An episode of joint attention was considered passive if both participants visually focused on an object for a minimum of 3 seconds but the child did not acknowledge the parent. 
If either participant looked away from the object for less than 3 seconds and then returned to the same object it was considered part of the same period of joint attention. 
Episodes joint attention were considered coordinated if  both participants visually focused on an object for a minimum of 3 seconds and at some point in the interaction the child indicated awareness of interaction with some overt behavior toward the parent. 
This could be looks to the parent's face, gestures, vocalizations, or turn-taking. 
If either participant looked away for less than 3 seconds and then returned to the same object, it was coded as part of the same period of joint attention. 
A second coder independently coded a third of the videos (i.e., 20 of the 60 videos, approximately equally distributed across ages) to establish reliability. % report reliability

# Results

## Lexical Diversity

## Joint Attention

# Experiment 2

@@

## Method

### Participants. 
84 infants aged 12-24 months and their parents participated in the same museum as Experiment 1. We included infants who were exposed to English at least 75 percent of the time or who were exposed less but whose participating parent reported that they primarily speak English with their child at home. [specific demo info]

### Materials. 
The design of Experiment 2 was similar to that of Experiment 1, except that instead of a No-Video control condition, parents instead watched a video that was generally related to child development research, but did not give any specific instructions about how to interact with infants or children. 
This was to control for the possibility that differences in language output and joint attention in Experiment 1 could be due to simply cuing parents to think about infants' learning and cognitive development. 
The videos presented in the Control Video condition were media clips (available on YouTube) of developmental psychologists explaining their research interleaved with footage of infants or toddlers engaged in developmental research studies. 
Thus, the content of the videos superficially matched those in the Activity Video condition, but did not suggest any particular activities.
The videos were trimmed to approximately match the average video length in the Activity Video condition (2 minutes ??).

### Procedure. 
The procedure for Experiment 2 matched that of Experiment 1, except that parents in the Control Video condition watched a control video before the play session. 
Consistent with the No-Video control condition in Experiment 1, parents in the Control Video condition were told to play with their child as they would at home, and were not given additional instructions.

# Results

## Lexical Diversity

## Joint Attention

# Discussion

For more information on citations in RMarkdown, see **[here](http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html#citations).**

## Footnotes

Indicate footnotes with a number\footnote{Sample of the first
footnote.} in the text. Place the footnotes in 9 point type at the
bottom of the page on which they appear. Precede the footnote with a
horizontal rule.\footnote{Sample of the second footnote.} You can also use 
markdown formatting to include footnotes using this syntax [^1].

[^1]: Sample of a markdown footnote.

## Two-column images

You can read local images using png package for example and plot 
it like a regular plot using grid.raster from the grid package. 
With this method you have full control of the size of your image. **Note: Image must be in .png file format for the readPNG function to work.**

You might want to display a wide figure across both columns. To do this, you change the `fig.env` chunk option to `figure*`. To align the image in the center of the page, set `fig.align` option to `center`. To format the width of your caption text, you set the `num.cols.cap` option to `2`.

```{r 2-col-image, fig.env = "figure*", fig.pos = "h", fig.width=4, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "This image spans both columns. And the caption text is limited to 0.8 of the width of the document."}
img <- png::readPNG("figs/walrus.png")
grid::grid.raster(img)
```

## One-column images

Single column is the default option, but if you want set it explicitly, set `fig.env` to `figure`. Notice that the `num.cols` option for the caption width is set to `1`.

```{r image, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "One column image."}
img <- png::readPNG("figs/lab_logo_stanford.png")
grid::grid.raster(img)
```

## R Plots

You can use R chunks directly to plot graphs. And you can use latex floats in the
fig.pos chunk option to have more control over the location of your plot on the page. For more information on latex placement specifiers see **[here](https://en.wikibooks.org/wiki/LaTeX/Floats,_Figures_and_Captions)**

```{r plot, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2, fig.height=2, fig.cap = "R plot" }
x <- 0:100
y <- 2 * (x + rnorm(length(x), sd = 3) + 3)

ggplot2::ggplot(data = data.frame(x, y), 
       aes(x = x, y = y)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

## Tables

@@

```{r xtable, results="asis"}
n <- 100
x <- rnorm(n)
y <- 2*x + rnorm(n)
out <- lm(y ~ x)

tab1 <- xtable::xtable(summary(out)$coef, digits=c(0, 2, 2, 1, 2), 
                      caption = "This table prints across one column.")

print(tab1, type="latex", comment = F, table.placement = "H")
```

# Acknowledgements

@@

# References

```{r}
# References will be generated automatically by Pandoc and included here.
# The following code is some latex to format the bibliography. Do not remove it.
```

@@

>>>>

% Template for Cogsci submission with R Markdown

% Stuff changed from original Markdown PLOS Template
\documentclass[10pt, letterpaper]{article}

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{float}
\usepackage{caption}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}

% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% hyperref package, useful for hyperlinks
\usepackage{hyperref}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% Sweave(-like)
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}
\DefineVerbatimEnvironment{Code}{Verbatim}{}
\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
\newenvironment{CodeChunk}{}{}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{apacite}

% KM added 1/4/18 to allow control of blind submission

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace}
%\doublespacing

% % Text layout
% \topmargin 0.0cm
% \oddsidemargin 0.5cm
% \evensidemargin 0.5cm
% \textwidth 16cm
% \textheight 21cm

\title{Understanding the impact of electronically-delivered parenting advice}

\author{{\large \bf Hanwen Vivian~Zhang (vivian3@stanford.edu)} \\ Department of Psychology, Stanford University \\ Palo Alto, CA 94301 USA \AND {\large \bf George~Kachergis (george.kachergis@gmail.com)} \\ \\ Department of Psychology, Stanford University \\ Palo Alto, CA 94301 USA \AND {\large \bf Michael C.~Frank (mcfrank@stanford.edu)} \\ \\ Department of Psychology, Stanford University \\ Palo Alto, CA 94301 USA}

\begin{document}

\maketitle

\begin{abstract}
Early parenting practices play an important role in shaping the future
outcomes of young children (Hart \& Risley, 1995; Heckman, 2006). In
particular, high quality early interactions and early language input
appear to facilitate more effective language learning and higher levels
of school performance. The rise in electronic parenting applications
(``apps'') holds the promise of delivering low-cost, positive
interventions on parenting style to a variety of different populations.
Of special interest are the parents of very young children, who are
often difficult to reach in other ways. Yet little is known about the
effects of communicating to parents through app-based interventions.
MTLD, types and tokens, joint attention, proportion of
nouns/verbs/adjectives, affect, concreteness, mean length of
utterance\ldots{}

\textbf{Keywords:}
digital parenting advice; joint action; lexical diversity; guided play;
\end{abstract}

\section{Introduction}\label{introduction}

\begin{itemize}
\tightlist
\item
  children learn through play -- physics, object features
\end{itemize}

\subsection{Guided Play Scaffolds
Learning}\label{guided-play-scaffolds-learning}

\begin{itemize}
\tightlist
\item
  guided play activities can scaffold children's language and reasoning
\end{itemize}

\subsection{Improving Parenting Practices and Language
Use}\label{improving-parenting-practices-and-language-use}

\begin{itemize}
\tightlist
\item
  early interventions on parenting practices can help children's
  outcomes
\end{itemize}

\subsection{Effectiveness of Digital
Delivery}\label{effectiveness-of-digital-delivery}

apps are an easy way to deliver curated, age-appropriate parenting
advice --- but can we measure an influence on parent's language? and do
the interventions lead to children paying more attention?

@@

\section{Experiment 1}\label{experiment-1}

@@

\subsection{Method}\label{method}

\subsubsection{Participants.}\label{participants.}

60 infants aged 6-24 months and their parents participated in a museum
in northern California. We included infants who were exposed to English
at least 50 percent of the time or who were exposed less but whose
participating parent reported that they primarily speak English with
their child at home. {[}include specific demo info here?{]} - good to
include now, and we can comment it out if it's too much (the full
report/paper will have it)

\subsubsection{Materials.}\label{materials.}

Stimuli included videos from a commercial parenting application, Kinedu
Inc. The videos were designed to show activities parents could perform
with their child to foster cognitive and physical development, and were
targeted to the child's age and development. In each video, an adult and
child perform the activity while a narrator explains the activity and
its purpose. We selected two videos for each of three age groups in our
sample (6-11.9 months, 12-17.9 months, 18-23.94 months). More
information about the specific videos is available in the Appendix.
Participants were also given a set of props corresponding to those in
the video they watched, so that they could complete the activity. The
props associated with each video are listed in the Appendix.

@@

\subsection{Procedure.}\label{procedure.}

After providing informed consent, parents in the Activity Video
condition watched the assigned video on a laptop with headphones. To
ensure that parents could give the video their full attention, the
experimenter played with the infant with a set of toys (different from
the experimental props used in the study) while the video was being
played. Immediately following the video, each parent-child dyad was
provided with the props to complete the activity they had viewed. The
toys were placed on a large foam mat, and parents were instructed to sit
on the mat with their child and re-create the activity they had viewed
for a period of three minutes. In the No-Video control condition, after
consenting parents were told to play with their child as they would at
home with the provided props for a period of three minutes. They were
not given any additional instructions about how to use the props.

@@

\subsubsection{Joint Attention Coding
Procedure.}\label{joint-attention-coding-procedure.}

Video of each session was coded for episodes of joint attention using
Datavyu software. The video taken at floor level was coded by default,
but the other video was referred to if the participants were not visible
or if there was a technological error with the first camera. For each
participant we coded for episodes of coordinated joint attention,
episodes of passive joint attention, and parental bids for joint
attention. We defined parental bids for joint attention as any attempt
to initiate joint attention (i.e label, point, or otherwise draw
attention to an object) that does not result in passive or coordinated
joint attention. If more than 3 seconds elapsed between bids, they were
coded as separate attempts. An episode of joint attention was considered
passive if both participants visually focused on an object for a minimum
of 3 seconds but the child did not acknowledge the parent. If either
participant looked away from the object for less than 3 seconds and then
returned to the same object it was considered part of the same period of
joint attention. Episodes joint attention were considered coordinated if
both participants visually focused on an object for a minimum of 3
seconds and at some point in the interaction the child indicated
awareness of interaction with some overt behavior toward the parent.
This could be looks to the parent's face, gestures, vocalizations, or
turn-taking. If either participant looked away for less than 3 seconds
and then returned to the same object, it was coded as part of the same
period of joint attention. A second coder independently coded a third of
the videos (i.e., 20 of the 60 videos, approximately equally distributed
across ages) to establish reliability. \% report reliability

\section{Results}\label{results}

\subsection{Lexical Diversity}\label{lexical-diversity}

\subsection{Joint Attention}\label{joint-attention}

\section{Experiment 2}\label{experiment-2}

@@

\subsection{Method}\label{method-1}

\subsubsection{Participants. 84 infants aged 12-24 months and their
parents participated in the same museum as Experiment 1. We included
infants who were exposed to English at least 75 percent of the time or
who were exposed less but whose participating parent reported that they
primarily speak English with their child at home. {[}specific demo
info{]}}\label{participants.-84-infants-aged-12-24-months-and-their-parents-participated-in-the-same-museum-as-experiment-1.-we-included-infants-who-were-exposed-to-english-at-least-75-percent-of-the-time-or-who-were-exposed-less-but-whose-participating-parent-reported-that-they-primarily-speak-english-with-their-child-at-home.-specific-demo-info}

\subsubsection{Materials.}\label{materials.-1}

The design of Experiment 2 was similar to that of Experiment 1, except
that instead of a No-Video control condition, parents instead watched a
video that was generally related to child development research, but did
not give any specific instructions about how to interact with infants or
children. This was to control for the possibility that differences in
language output and joint attention in Experiment 1 could be due to
simply cuing parents to think about infants' learning and cognitive
development. The videos presented in the Control Video condition were
media clips (available on YouTube) of developmental psychologists
explaining their research interleaved with footage of infants or
toddlers engaged in developmental research studies. Thus, the content of
the videos superficially matched those in the Activity Video condition,
but did not suggest any particular activities. The videos were trimmed
to approximately match the average video length in the Activity Video
condition (2 minutes ??).

\subsubsection{Procedure.}\label{procedure.-1}

The procedure for Experiment 2 matched that of Experiment 1, except that
parents in the Control Video condition watched a control video before
the play session. Consistent with the No-Video control condition in
Experiment 1, parents in the Control Video condition were told to play
with their child as they would at home, and were not given additional
instructions.

\section{Results}\label{results-1}

\subsection{Lexical Diversity}\label{lexical-diversity-1}

\subsection{Joint Attention}\label{joint-attention-1}

\section{Discussion}\label{discussion}

For more information on citations in RMarkdown, see
\textbf{\href{http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html\#citations}{here}.}

\subsection{Footnotes}\label{footnotes}

Indicate footnotes with a number\footnote{Sample of the first
footnote.} in the text. Place the footnotes in 9 point type at the
bottom of the page on which they appear. Precede the footnote with a
horizontal rule.\footnote{Sample of the second footnote.} You can also
use markdown formatting to include footnotes using this
syntax.\footnote{Sample of a markdown footnote.}

\subsection{Two-column images}\label{two-column-images}

You can read local images using png package for example and plot it like
a regular plot using grid.raster from the grid package. With this method
you have full control of the size of your image. \textbf{Note: Image
must be in .png file format for the readPNG function to work.}

You might want to display a wide figure across both columns. To do this,
you change the \texttt{fig.env} chunk option to \texttt{figure*}. To
align the image in the center of the page, set \texttt{fig.align} option
to \texttt{center}. To format the width of your caption text, you set
the \texttt{num.cols.cap} option to \texttt{2}.

\begin{CodeChunk}
\begin{figure*}[h]

{\centering \includegraphics{figs/2-col-image-1}

}

\caption[This image spans both columns]{This image spans both columns. And the caption text is limited to 0.8 of the width of the document.}\label{fig:2-col-image}
\end{figure*}
\end{CodeChunk}

\subsection{One-column images}\label{one-column-images}

Single column is the default option, but if you want set it explicitly,
set \texttt{fig.env} to \texttt{figure}. Notice that the
\texttt{num.cols} option for the caption width is set to \texttt{1}.

\begin{CodeChunk}
\begin{figure}[H]

{\centering \includegraphics{figs/image-1}

}

\caption[One column image]{One column image.}\label{fig:image}
\end{figure}
\end{CodeChunk}

\subsection{R Plots}\label{r-plots}

You can use R chunks directly to plot graphs. And you can use latex
floats in the fig.pos chunk option to have more control over the
location of your plot on the page. For more information on latex
placement specifiers see
\textbf{\href{https://en.wikibooks.org/wiki/LaTeX/Floats,_Figures_and_Captions}{here}}

\begin{CodeChunk}
\begin{figure}[H]

{\centering \includegraphics{figs/plot-1}

}

\caption[R plot]{R plot}\label{fig:plot}
\end{figure}
\end{CodeChunk}

\subsection{Tables}\label{tables}

@@

\begin{table}[H]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & t value & Pr($>$$|$t$|$) \\ 
  \hline
(Intercept) & 0.04 & 0.11 & 0.3 & 0.74 \\ 
  x & 2.05 & 0.11 & 18.0 & 0.00 \\ 
   \hline
\end{tabular}
\caption{This table prints across one column.} 
\end{table}

\section{Acknowledgements}\label{acknowledgements}

@@

\section{References}\label{references}

@@

\hypertarget{refs}{}
\hypertarget{ref-Hart1995}{}
Hart, B., \& Risley, T. R. (1995). \emph{Meaningful differences in the
everyday experience of young american children}. Baltimore, MD: Brookes.

\hypertarget{ref-Heckman2006}{}
Heckman, J. J. (2006). Skill formation and the economics of investing in
disadvantaged children. \emph{Science}, \emph{312}(5782), 1900--1902.
\url{http://doi.org/10.1126/science.1128898}

\bibliographystyle{apacite}

\end{document}
